\begin{thebibliography}{1}

\bibitem{sj:jasa85}
H.~Schmidt and F.B. Jensen.
\newblock A full wave solution for propagation in multilayered viscoelastic
  media with application to {G}aussian beam reflection at fluid-solid
  interfaces.
\newblock {\em J. Acoust. Soc. Am.}, 77:813--825, 1985.

\bibitem{st:gjras86}
H.~Schmidt and G.~Tango.
\newblock Efficient global matrix approach to the computation of synthetic
  seismograms.
\newblock {\em Geophys. J. R. Astr. Soc.}, 84:331--359, 1986.

\bibitem{jkps}
F.B. Jensen, W.A. Kuperman, M.B. Porter, and H.~Schmidt.
\newblock {\em Computational Ocean Acoustics}.
\newblock American Institute of Physics, New York, 1994.

\bibitem{hs:saf}
H.~Schmidt.
\newblock {S}{A}{F}{A}{R}{I}: Seismo-acoustic fast field algorithm for range
  independent environments. user's guide.
\newblock SR 113, SACLANT ASW Research Centre, La Spezia, Italy, 1987.

\bibitem{stern}
M.~Stern, A.~bedford and H.R.~Millwater.
\newblock Wave reflection from a sediment layer with depth-dependent properties.
\newblock  {\em j. Acoust.Soc. Am.}, 77:1781--1788, 1985.

\bibitem{hovem}
J.M.~Hovem and J.D.~Ingram.
\newblock Viscous attenuation of sound in saturated sand.
\newblock  {\em J. Acoust. Soc. Am.}, 66:1807--1812, 1979.

\bibitem{yavari}
B.~Yavari and A.~Bedford.
\newblock Comparison of numerical calculations of two Biot
coefficients with analytical solutions.
\newblock  {\em J. Acoust. Soc. Am.}, 90:985--990, 1991.

\bibitem{biot}
M.A.~Biot.
\newblock Theory of propagation of elastic waves in a fluid saturated
porous solid.
\newblock  {\em J. Acoust. Soc. Am.}, 28:168--191, 1956.

\bibitem{sk:jasa94}
H. Schmidt and W.A. Kuperman, Spectral and modal representations of
the Doppler-shifted field in ocean waveguides, 
\newblock {\em J. Acoust. Soc. Am.}, 96(1):386--395, 1994.

\bibitem{jkim}
J.S. Kim.
\newblock {\em Radiation from directional seismic sources in laterally
  stratified media with application to {A}rctic ice cracking noise}.
\newblock PhD thesis, Massachusetts Institute of Technology, May 1989.

\bibitem{sg:jasa85}
H.~Schmidt and J.~Glattetre.
\newblock A fast field model for three-dimensional wave propagation in
  stratified environments based on the global matrix method.
\newblock {\em J. Acoust. Soc. Am.}, 78:2105--2114, 1985.

\bibitem{sk:jasa88}
H.~Schmidt and W.A.~Kuperman.
\newblock Estimation of surface noise source level from low-frequency
seismoacoustic ambient noise measurements. 
\newblock  {\em J. Acoust. Soc. Am.}, 84:2153--2162, 1988.

\bibitem{bks:jasa88}
A.B.~Baggeroer, W.A.~Kuperman and H.~Schmidt.
\newblock Matched field processing: Source localization in correlated
noise as an optimum parameter estimation problem.
\newblock  {\em J. Acoust. Soc. Am.}, 83:571--587, 1988.

\bibitem{sbk:jasa90}
H. Schmidt, A.B.~Baggeroer, W.A.~Kuperman and E.K. Scheer. 
\newblock Environmentally tolerant beamforming for high-resolution
matched field processing: deterministic mismatch.
\newblock  {\em J. Acoust. Soc. Am.}, 88:1851--1862, 1990.

\bibitem{sk:rvb94}
H.~Schmidt and W.A.~Kuperman.
\newblock Spectral representations of rough interface reverberation in stratified ocean waveguides,
\newblock  {\em J. Acoust. Soc. Am.}, 96:Submitted, 1994.

\bibitem{ks:jasa89}
W.A. Kuperman and H. Schmidt,
\newblock Self-consistent perturbation approach to rough surface scattering in stratified elastic media,
\newblock  {\em J. Acoust. Soc. Am.}, 86:1511--1522, 1989.

\bibitem{nato_bs_95}
 A.B.Baggeroer and H. Schmidt, ``Parameter estimation bounds and
the accuracy of full field inversions.'' In O. Diachok, et al, editor,
{\em Full field inversion methods in ocean and seismic acoustics},
Kluwer Academic Publishers, Dordrecht, The Netherlands, 1994.

\bibitem{nato_sb_95}
H. Schmidt and A.B.Baggeroer.  ``Physics-imposed resolution and
robustness issues in seismo-acoustic parameter inversion''.  In
O.Diachok, et al, editor, {\em Full field inversion methods in ocean
and seismic acoustics}, Kluwer Academic Publishers, Dordrecht, The
Netherlands, 1994.

\bibitem{Schmidt_95}
 H. Schmidt, W. Seong, and J.T. Goh, ``Spectral super-element
approach to range-dependent ocean acoustic modeling,'' {\em J. Acoust.
Soc. Am.}, 98(1), 465--472, 1995.


\bibitem{Goh_96} 
J.T. Goh and H. Schmidt, A hybrid coupled wavenumber integration
approach to range-dependent seismo-acoustic modeling,
J. Acoust. Soc. Am, 100, 1409--1420, 1996.

\bibitem{Goh_97}
J.T. Goh, H. Schmidt, P. Gerstoft and W. Seong, ``Benchmarks
for validating range-dependent seismo-acoustic propagation codes'', IEEE
J. Oceanic Eng., April 1997.

\bibitem{Stoll_81}
R.D. Stoll and T.K. Kan, ``Reflection of acoustic waves at a water
sediment interface,'' J. Acoust. Soc. Am., 70, 149--156, 1981.

\bibitem{Saclant94}
H. Schmidt and A.B. Baggeroer, "Physics-imposed resolution and robustness 
        issues in seismo-acoustic parameter inversion", 
        In "O. Diachok {\em et al.} Eds.'' {\em Full field inversion methods 
                in ocean and seismic
		acoustics}, Kluwer Academic Publishers,
	        Dordrecht, The Netherlands, 1994.


\bibitem{Lepage_99} 
K. Lepage and H. Schmidt,
        "Spectral integral representations of scattering from
                  volume inhomogeneities in ocean waveguides", 
Submitted to J. Acoust. Soc. Am., June 1999.

\bibitem{Turgut_97} 
A. Turgut, ``Inversion of bottom/subbottom statistical parameters from acoustic backscatter data'', J. Acoust. Soc. Am., 102(2), 833--852, 1997.
  
\end{thebibliography}



